### 2023-03-06 theory.tex
MJ first edits

\documentclass[11pt]{article}
\setlength{\parindent}{10pt}
\usepackage [square,colon]{natbib}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage {comment}
\usepackage[flushleft]{threeparttable}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=1.5cm]{geometry}
\usepackage{doi}
\usepackage{subcaption}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{marginnote}
\usepackage[T1, OT1]{fontenc}
\usepackage{wrapfig}
\linenumbers
\errorcontextlines=3
\usepackage{hyperref}
\usepackage{lscape}
\linenumbers
\usepackage[explicit]{titlesec}
\usepackage{setspace}
\usepackage{xcolor}   % Make the colors used for the URL available
\hypersetup{ % Remove ugly boxes around URLs
  colorlinks, linkcolor={black!50!black}, citecolor={black!50!black}, urlcolor={black!80!black}
}

\usepackage[export]{adjustbox}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

\newenvironment{laysummary}{
  \chapter{\centering \textbf{Plain Language Summary}}
}

\newenvironment{abstract_}{
  \chapter{\centering \textbf{Abstract}}
}

\author[1]{Ian Delaney}
\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne, B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
\title{Comparison of sediment transport capacity in subglacial and fluvial systems }

\begin{document}
\maketitle

\paragraph{Key findings:}
\begin{enumerate}
\item Sediment transport capacity varies more in pressurized subglacial channels compared to open fluvial ones, in response to changing water discharge due to the fixed size of their channels.  I can't fully undertand this. Too much information is overlain here. Separate concepts and give a more linear explanation. 
\item  As water discharge changes, differences in the sediment transport capacity response of fluvial vs subglacial channels  could substantially  impact ________ what kind? processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments (?). 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends largely on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
In fluvial channels, changing water discharge can be accommodated by changes in? water depth, width of water flow [ is this width of water flow? ], and water velocity.
In subglacial channels, however, water flows through a conduit with a fixed size and shape?, and the water is pressurized by the overlying glacier ice.
As a result, changing water discharge may result in modifying the water velocity, but not the  shape of the channel.
Parameterization of _______ indicates that a calculated? observed? increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.  This last sentence seems repetitive. I'm not fully understanding the results. Suggest revise after a complete update of the manuscript.

\vspace{0.5cm}

\laysummary % 200 words
It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and fluvial channels.
In both subglacial and fluvial channels, sediment transport capacity depends, to the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
The shear stress depends largely on water velocity.
In fluvial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
As a result, sediment transport capacity varies far more in subglacial channels compared to fluvial ones, even though changing channel width can accommodate some sediment transport variability in fluvial channels.
This increased variability has several important implications for sediment transport in glacial systems compared to fluvial ones.
This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.

%% TC:endignore
\begin{spacing}{1}
  \section{Introduction}

  Glaciers expel large quantities of sediment  from their termini \citep{hallet1996}. NEEDS MORE REFERENCES! Current 
  changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions, such as ________, _________, _____ \citep[e.g.][]{zhang2022}.    NEEDS MORE REFERENCES to fit with geographic sites!     Get the observations and measurements well stated, first. 
  Numerical models integrated with glacial monitoring data  consider evaluate?  the subglacial component of sediment discharge ___________at glacial termini? as a function of the availability of sediment from  erosion of bedrock or  prior existence of material and the sediment transport capacity generated by fluvial processes below the glacier ice? \citep{creyts2013,brinkerhoff2017,beaud2018,delaney2019}.
  These models  essentially represent subglacial sediment transport processes as similar to sediment transport processes in rivers, with the main difference being that when water flows through a subglacial conduit, it is pressurized by the ice above \citep{rothlisberger1972}.

 Shear stress between the water and the sediment that it flows over has a first-order control on the transport of sediment in both  subglacialand fluvial systems, \citep{shields1936} , but only if an adequate quantity of sediment is present so as to be transported fluvially (i.e. in a ``transport-limited'' regime). Confusing ; fluvial is used in too many contexts. How can you clarify this?
 between the water and the sediment that it flows over : this is an awkward phrasing. Is this correct: shear stress applied by the flowing water on the channel bed?
  The shear stress increases with increasing water velocity and roughness of the channel.
  Greater shear stress will mobilize increased quantities of sediment, along with larger sediment clasts. Confusing - need to specify grain size here. what is meant by clast?
  Sediment transport relationships described by  \citet{shields1936}, \citet{meyer1948},  and \citet{engelund1967} , for example, use shear stress, or an approximation thereof [ is this needed here?, confusing], to evaluate sediment transport capacity per unit width of ______?.
  The total sediment transport capacity across a channel's cross section depends on the integrated shear stress over the entire channel bed (Figure~\ref{fig:cartoon}). Here, the channel cross section refers to _______ (site a, Fig. ___) and the channel bed refers to ____ (site b, Fig. __)
Thus,  channel geometry plays a key role in sediment transport in both fluvial channels and subglacial ones because _________________ state the reason why \citep[e.g.][]{church2006,delaney2019}.

In a regime that is not transport limited [??? How is this different from the previous paragraph?]  the behavior of _________ what process? in subglacial channels differs from that in fluvial channels  in response to changing water discharge, Again, not following here... is there changing water discharge in the paragraph above?

  In subglacial environments? systems? , the water is pressurized and the changing? size of thechannel responds to the roughness of the walls and the water velocity   \citep[Figure~\ref{fig:cartoon}\,a and c; e.g. ][]{rothlisberger1972}. you have used "channel" many times in this sentence - confusing.
 Diurnal variations in water flux may cause water to flow through a conduit whose size does not respond at the same timescale(s) as the increase in water flux \citep{rothlisberger1972}.
This means that increased water discharge in subglacial channels can result in greater water velocity and thus sediment transport capacity.
  At the same time, a relatively  fixed channel width  means that the distance over sediment is mobilized does not evolve over the same timescale(s) as water discharge \citep[e.g.][]{werder2010}.

In fluvial systems with open-channel flow [why is this specified as compared with subglacial system? needed?], a commensurate [relative to ___?, delete?] increase in water discharge will be accommodated by  a reduced increase [whoa - hard to figure out] a reduction? in water velocity along with an increase in channel width and/or depth (Figure~\ref{fig:cartoon}\,b and d).
  The change in water velocity could be lower [not following here], but the width over which sediment can be mobilized changes \citep{leopold1953}.  [not following here, think through a clear explanation even if this requires more sentences. ]
  Thus, the systematic response in sediment transport capacity with respect to changing water discharge is  different in the two systems because ______________.

 Differences in the processes of ________________in subglacial and fluvial systems have been implicitly included in a wide range of available models that evaluate sediment transport relationships.  \citep[e.g.][]{walder1994,tucker1997,beaud2018,wickert2019}.
  Furthermore, models of \textbf{fluvial} [again, two uses of fluvial in this sentence is confusing] transport of sediment in glacial and \textbf{fluvial} systems investigate the relationship between sediment dynamics and climate change? \citep[e.g.][]{tucker1997,delaney2020}
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and fluvial systems has been less well discussed.

 This manuscript leverages parameterizations of subglacial hydraulics and fluvial hydraulics to explore  differences in the relationships among sediment transport capacity and water discharge. It also explores  the impact of water discharge variability on sediment transport capacity in idealized? subglacial and fluvial  two systems.
  Lastly, the implications for sediment transport processes in subglacial systems, compared to fluvial ones, are discussed.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water? Velocity magnitude in the glacial, $v_g$ , and fluvial (river?) ______? ,$v_f$ ,systems are shown by arrow length and lines of ______greatest value? marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial (river?) channels.  Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methods}
  \label{sect:meth}
 Relationships among water discharge, subglacial water velocity, and channel geometry in both fluvial (river?) and subglacial channels (Table ___) are described with subglacial and fluvial(word?) parameterization models.
  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculatethis value and integrate  it across the channel bed (Fig. __).
  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936}. and an eventual sediment transport capacity.  [is this what you mean?]

  \ian{check units on fluvial channel factor}
  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (fluvial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (fluvial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
      Fluvial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used. Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the? channel  and geometry shape and size? will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  It is also assumed?? More, specifically,  the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
  smaller values of $\beta$ result in shallow, wide channels. 
  The width of the flat channel floor $w_c$ , where beta equals _____? is given by
  \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge; it is not used ?? to evaluate the hydraulic gradient of the channel in Equation~\ref{eq:DW}.
  It is assumed that?? the channel size is usually not in equilibrium with the quantity of water flowing through it.
  With this formulation, the processes in an R-channel can be represented algebraically \citep{delaney2019}. What is an "R-channel"? doesn't seem to be defined above.

  For purposes discussed below, we evaluate the width integrated shear stress as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Fluvial channel  parameterization}
  \label{sect:fluv}

 To parameterize?? describe? the shear stress of water flowing across sediments in the  fluvial channel  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
  Here, it is assumed that there is  conservation of mass and a sufficiently wide channel so that the hydraulic radius, ____,  is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for fluvial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  %
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
 \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  %
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

  \subsection{Implementation}

  The formulation for the _______ of the glacial system , eqn _____, requires inputs of the hydraulic gradient, ___, water discharge, ____, representative water discharge, ____,  Hooke angle $\beta$ \citep{hooke1990}, and friction factor, $f_r$. Help the reader find these quantities and equations!
  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity, ___________ m/s, in accordance with  measurements made through dye-tracing \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}. in xyz glacial environments.  Give the reader context for this!

  In the fluvial river? system, the water is not pressurized, so only the bed slope is needed to define the _____________, eqn ______. The friction factor, $f_f$ ,and channel shape factor, $k$, are assigned (Section~\ref{sect:fluv}), using xyz measurements from abc stream? river? environments.

  The Reynolds number $R_e$, which gives an evaluation of the ________________, is given as 
  \begin{equation}
    \label{eq:re}
    R_e\,=\, v \,\frac{D_h}{\nu},
  \end{equation}
  % 
  \noindent where $\nu$ is the kinematic viscosity of water, and  $v$ is velocity of water, $v_g$ or $v_f$.

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating __________________.

  \section{Results}
  First, the subglacial hydraulics parameterization is applied to a representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
  Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge. The fluvial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the fluvial system is in green.  Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the fluvial system (Figure~\ref{fig:model_outs}).
  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the fluvial system the water velocity only varies by $0.11$.
  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the fluvial system varies by roughly $0.21$. This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.

  Changing channel geometry accommodates evolving water discharge in the fluvial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
  The range of values between width-integrated shear stress in the fluvial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
  Indeed, the relative  difference in width-integrated shear stress is less between the fluvial and glacier cases, compared to the water velocity and shear stress outputs.
  The reduced variability in width-integrated shear stress results from the additional width across the fluvial channel that accommodates a portion of the sediment transport capacity.

  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma). $Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (fluvial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to fluvial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to fluvial one (Figure~S1).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to fluvial ones.
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the fluvial system.
  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to fluvial ones (Figure~\ref{fig:range} and S1).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the fluvial one (Figure S1).

  \section{Discussion}
  \subsection{Assumptions}

  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for $\beta$ and $k$.

  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
  While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.

  In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might slightly modify the width of the subglacial channel.

  Both subglacial and fluvial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous (Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}). 
  The fluvial parameterization assumes that sediment is transported across all channel widths, as opposed to bankfull width when much sediment transport occurs \citep{wolman1960}.
  As a result, the variability in the fluvial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).

  \subsection{Geomorphic Implications}
  \label{sect:GI}
  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the fluvial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to fluvial ones.
  Shear stress is scaled to the $\frac{3}{2}$ power  and relies on shear stress exceeding a threshold in sediment transport relationships  such as \citet{meyer1948}.
  In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
  In both cases, the exponent greater than $1$ magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).

  \subsubsection{Sediment transport below glaciers and at their margins}

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment more frequently may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime \citep[e.g.][]{herman2015}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small conduit, not conditioned by the amount of water flow, will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).

  Increased variability in sediment transport also impacts sediment export from glacier margins into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
  Here, in the transition to open-channel flow as the water leaves the glacier, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), potentially resulting in sediment deposition once sediment enters the proglacial area.
  Note that the transition from pressurized to open channel flow could occur above the glacier's terminus, when cavity closure rates are reduced and water can flow freely \citep{egli2021b}, resulting in intermittent deposition and remobilization of sediment below the glacier terminus \citep{perolo2018}.

  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many fluvial systems compared to subglacial ones.
  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:range}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
  Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
  Thus the disparity in sediment transport capacity between the subglacial and fluvial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).

  \subsubsection{Records of sediment discharge  from glaciers}

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}, in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
  Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in fluvial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}, but could be the result of reduced conduit size and therefore increased sediment transport capacity.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to fluvial systems \citep{jansson2005,ganti2016}. 
  For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity, in addition to water discharge \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972}.
  Conversely, sediment transport in most fluvial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
  Furthermore, if the glacier is in a supply-limited case, the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.

  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to fluvial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to fluvial systems, to overcome the increased variability in sediment transport capacity.

  \section{Conclusions}

  Application of parameterizations for subglacial and fluvial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
  Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in fluvial ones.
  The variability in shear stress is reduced when the evolving channel width is accounted for in fluvial systems.
  Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to fluvial ones.

  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple fluvial processes together.
  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
  The inconsistent response of sediment transport in subglacial and fluvial systems to changing water discharge could add noise to sediment transport signals between the systems.
  In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone. 
  The differences in fluvial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.

  \section{Acknowledgments}

  I was funded by SNSF Project No. PZ00P2\_202024.
  The julia code used herein is included in the supplementary material.

\end{spacing}

\bibliographystyle{apalike}
\bibliography{Paperlib.bib}

\section{Supplement}

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and fluvial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

% LocalWords:  pre todo phillips CVODE decadal refreezing Mauro eq%  LocalWords:  dQs
% LocalWords:  dx ms ca DF Hm meltwater Griessee gridded ehrbar ng
% LocalWords:  koppes costa bendixen delaney herman hallet creyts ian
% LocalWords:  farinotti stott rickenman delaney2019 plagerized
% LocalWords:  fischer zekollari clarke raymond larour delaneyinprep
% LocalWords:  guillon exner wolman engelund Weisbach

As a result, the parameterization here is most relevant for relatively wide rivers, where channel width evolves greatly with water discharge (Equation~\ref{eq:wcf}).
% This is probably the case in many proglacial rivers and glacierized catchments where the divergent behavior of subglacial and fluvial channels is most relevant (Figure~\ref{sect:GI}) \ian{Find reference about the geometry of proglacial river}.

Qbs_out Qws_out