### 2023-03-07 theory.tex
polishing

\documentclass[11pt]{article}
\setlength{\parindent}{10pt}
\usepackage [square,colon]{natbib}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage {comment}
\usepackage[flushleft]{threeparttable}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=1.5cm]{geometry}
\usepackage{doi}
\usepackage{subcaption}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{marginnote}
\usepackage[T1, OT1]{fontenc}
\usepackage{wrapfig}
\linenumbers
\errorcontextlines=3
\usepackage{hyperref}
\usepackage{lscape}
\linenumbers
\usepackage[explicit]{titlesec}
\usepackage{setspace}
\usepackage{xcolor}   % Make the colors used for the URL available
\hypersetup{ % Remove ugly boxes around URLs
  colorlinks, linkcolor={black!50!black}, citecolor={black!50!black}, urlcolor={black!80!black}
}

\usepackage[export]{adjustbox}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\marie}[1]{{\textbf{\color{green}Marie says:} \color{green} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

\newenvironment{laysummary}{
  \chapter{\centering \textbf{Plain Language Summary}}
}

\newenvironment{abstract_}{
  \chapter{\cemuchntering \textbf{Abstract}}
}

\author[1]{Ian Delaney}
\affil[1]{Institut des dynamiques de la surface terrestre (IDYST), Universit\'{e} de Lausanne

 B\^{a}timent G\'{e}opolis, CH-1015 Lausanne}
\title{Comparison of sediment transport capacity variations in subglacial and subaerial systems }

\begin{document}
\maketitle

\paragraph{Key findings:}
\begin{enumerate}
\item The sediment carrying capacity of subglacial water varies more in response to changing water discharge compared to subaerial channels, as subglacial channels are pressurized by the ice above.
\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel.
%\marie{I can't fully understand this. Too much information is overlain here. Separate concepts and give a more linear explanation.} 
\item  As water discharge changes, differences in the sediment transport capacity response of subaerial vs subglacial channels  could substantially  impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments. 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends mainly on the velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
In subglacial channels, however, water flows through a conduit with a fixed size and shape, and the water is pressurized by the overlying glacier ice.
As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
Parameterization of channel dynamics indicates that increased variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
The increased variability in sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.  %\marie{This last sentence seems repetitive. I do not fully understand the results. Suggest revision after a complete update of the manuscript}

\vspace{0.5cm}

\laysummary % 200 words
It is well known that sediment transport capacity responds differently to changing water discharge in subglacial and subaerial channels.
In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over-which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom.
The shear stress depends largely on water velocity.
In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
As a result, sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
This increased variability has several important implications for sediment transport in glacial systems compared to subaerial ones.
%\maire{This is hard to follow. Suggest you rewrite after greater clarity in the rest of the paper.}

%% TC:endignore
\begin{spacing}{1}
  \section{Introduction}

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Increased sediment discharge at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.

Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson1993,herman2021}. 
 On shorter time periods, pressurized water transports available sediment from the glacier \citep{creyts2013,beaud2018b,delaney2019}, should enough sediment be present below the glacier (i.e. in a "transport"-limited regime).

In a "transport"-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
\begin{linenomath*}
    \begin{equation}
      \label{eq:tau_t}
        \tau_t \propto w\, v^2,
        \end{equation}
  \end{linenomath*}
  % 
  where $w$ is the width of the channel.
  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:v}
  v = \frac{Q_w}{S(w)},
  \end{equation}
  \end{linenomath*}
%
where $S$ is the channel's cross-sectional area that responds to channel width $w$, and $Q_w$ is water discharge.

In subaerial channels, such as rivers with open channel flow, the depth $d$ and width $w$ of the channel evolve with changing water discharge $Q_w$ \citep{leopold1953}. During a flood with greater water discharge $Q_w$, for instance, cross-sectional area $S$ increases with channel width $w$ along with velocity $v$. As a result, the shear stress across the channel $\tau_t$ will increase in response to both width $w$ and $v$.

Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}. As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
Changes in water discharge $Q_w$ are mainly accommodated in water velocity $v$, because the width $w$ and depth $d$ of the channel are fixed (Figure~\ref{fig:cartoon}).
It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to primarily respond to the water velocity $v$, as opposed to the subaerial system where width $w$ may evolve as well.

As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,delaney2019}.
  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003}.
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.

 This manuscript leverages parameterizations of subglacial and subaerial hydraulics to explore differences in the relationships between  water discharge, channel geometry, and water velocity. 
 With these parameterizations, the manuscript discusses the impact of water discharge variability on sediment transport capacity in idealized subglacial and subaerial systems.
  Lastly, the implications for sediment transport processes in subglacial systems, compared to subaerial ones, are discussed.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length \marie{and lines of xxx greatest value} marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and subaerial channels. Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methods}
  \label{sect:meth}
 Relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}) are described with subglacial and subaerial parameterizations.
  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig. \ref{fig:cartoon}).
  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep{shields1936} and thus the calculation of sediment transport capacity.

  \ian{check units on subaerial channel factor}
  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, subaerial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

      \textbf{Parameters and Constants}  & & &\\
      Gravitational constant&$g$& $-9.81$&$\mathrm{m\,s^{-2}}$\\
      Density of water & $\rho_w$& $1000$ & $\mathrm{kg\,m^{-3}}$ \\
      Density of ice & $\rho_i$& $900$ & $\mathrm{kg\,m^{-3}}$ \\
      Kinematic viscosity of water &$\nu$& $10^{-4}$& $\mathrm{m^2\,s^{-1}}$\\
      Hooke angle of channel & $\beta$ & $\frac{\pi}{2}$ & \unit{rad}\\
      Flotation fraction & $f_f$&$0.7$& $\mathrm{(-)}$\\
      Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
      Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
      Gradient of glacier surface & $\frac{\partial z_{sg}}{\partial x}$ &$0.25$& $\mathrm{(-)}$\\
      Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
      Subaerial channel factor & $k$ &$3$ & $\mathrm{m^{-2}\, s}$\\
      Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
      \hline
    \end{tabular}
    \label{table:vpm}
  \end{table}

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used. Here, it is assumed that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Thus, the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$ such that:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

  In Equation~\ref{eq:DW}, $f_r$ is the Darcy-Weisbach friction factor and $\rho_w$ is the density of water (Table \ref{table:vpm}). $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
  \begin{equation}
    \label{eq:Hf}
    s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
  \end{equation}
  where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
  smaller values of $\beta$ result in shallow, wide channels. 
  The width of the flat channel floor $w_c$ , given $beta$ is   \begin{equation}
    \label{eq:dh2wc}
    w_{cg} = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S}{\beta -\sin \beta}},
  \end{equation}
  % 
  where $S$ is the cross-sectional area of the channel given by
  \begin{equation}
    \label{eq:dh2S}
    S_g =  \frac{D_h^2}{2}~ \frac{\Big(\frac{\beta}{2} \,+ \, \sin \frac{\beta}{2}\Big)^2  }{\beta\,-\,\sin \beta}.
  \end{equation}

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \citep{rothlisberger1972,delaney2019}.

  For purposes discussed below, width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

  \subsection{Subaerial channel  parameterization}
  \label{sect:fluv}

 To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
  Here, it is assumed that there is conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  %
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
 \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  %
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

  \subsection{Implementation}

The formulation of the glacial system (Section~\ref{sect:sub_mode}) requires inputs of the hydraulic gradient (Equation \ref{eq:psi}; $\frac{\partial z_{sg}}{\partial x}$, $\frac{\partial z_{c}}{\partial x}$), water discharge, $Q_w$ representative water discharge, $Q_w^*$,  Hooke angle (Equation~\ref{eq:Hf}; $\beta$), and friction factor (Equation~\ref{eq:DW}; $f_r$). \marie{Help the reader find these quantities and equations}
  Values of $\beta$ and $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.

  In the subaerial system, the water is not pressurized, so only the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}. The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).

  The Reynolds number $R_e$, which gives an evaluation of the degree of turbulence of the water, is given as 
  \begin{equation}
    \label{eq:re}
    R_e\,=\, v \,\frac{D_h}{\nu},
  \end{equation}
  % 
  \noindent where $\nu$ is the kinematic viscosity of water, and  $v$ is velocity of water, $v_g$ or $v_f$.

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.

  \section{Results}
  First, the subglacial hydraulics parameterization is applied to a representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
  Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge. The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the subaerial system is in green.  Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$.
  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$. This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.

  Changing channel geometry accommodates evolving water discharge in the subaerial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
  The range of values between width-integrated shear stress in the subaerial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs.
  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates a portion of the sediment transport capacity.

  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma). $Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and subaerial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (subaerial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S1).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones.
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the subaerial system.
  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure S1).

  \section{Discussion}
  \subsection{Assumptions}

  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for channel shape factors $\beta$ and $k$.

  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
  While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.

  In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might slightly modify the width of the subglacial channel.

  Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}][]{wainwright2015}. 
  The subaerial parameterization assumes that sediment is transported across all channel widths, as opposed to large flows such as floods when much sediment transport occurs \citep{wolman1960}.
  As a result, the variability in the subaerial system across the range of discharge values could be underestimated in the parameterization (Figure~\ref{fig:range}).

  \subsection{Geomorphic Implications}
  \label{sect:GI}
  The consistent relationship between water discharge variability and relative variability of the water velocity, shear stress, and width-integrated shear stress results from the fixed geometry in the glacier system and variable geometry in the subaerial system as they respond to changes in water discharge (Equations~\ref{eq:tau} and \ref{eq:Sf}).
  Increased variability in sediment transport capacity with respect to discharge has several implications for sediment transport in subglacial systems compared to subaerial ones.
  Shear stress is scaled to the $\frac{3}{2}$ power and relies on shear stress exceeding a threshold in sediment transport relationships  such as in \citet{meyer1948}.
  In the total sediment transport relationship \citet{engelund1967}, the shear stress is scaled to $\frac{5}{2}$ power.
  In both cases, the exponent greater than $1$ of shear stress in the sediment transport relationship magnifies the variability in sediment transport beyond the highly variable shear stress and width-integrated shear stress described above in subglacial environments (Figure~\ref{fig:model_outs}\, b and d; Figure~S1).

  \subsubsection{Sediment transport below glaciers and at their margins}

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a very large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).

  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
  Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier. The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
  Note that the transition from pressurized to open channel flow could occur above glacier termini when cavity closure rates are reduced and water can flow freely \citep{egli2021b}. 
  Periodic sediment deposition in these areas may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.

  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{costa2017,vanas2017,delaney2018,hasholt2018}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.
  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability from subglacial to proglacial systems (Figure~\ref{fig:range}\, d,\,e,\,f), while the variability in the subglacial system remains higher (Figure~S1).
  Therefore, at locations such as the Greenland Ice Sheet, with smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{delaney2018,hasholt2018}, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic than alpine systems as water leaves the glacier and moves downstream.
  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).

  \subsubsection{Records of sediment discharge  from glaciers}

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}, in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
  Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size and therefore increased sediment transport capacity.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
  For instance, if a glacier expels sediment in a transport-limited regime, the pressurized nature of subglacial water flow means sediment discharge capacity responds to the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity, in addition to water discharge \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972}.
  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006}. 
  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of sliding  \citep{herman2015,seguinot2021}, adding additional processes represented in the sediment discharge signal \citep{delaney2019}.

  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, given the potential amount of additional noise \citep{castletort2003,jerolmack2010,romans2016}.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.

  \section{Conclusions}

  \ian{simplify as did in introduction}

  Application of parameterizations for subglacial and subaerial water flow shows that variable processes occur in sediment transport capacity's response to water discharge, due to their fixed and variable channel geometry, respectively.
  Results show that the impact of variable water discharge on water velocity and shear stress, first-order controls on sediment mobilization, is much higher in subglacial channels than in subaerial ones.
  The variability in shear stress is reduced when the evolving channel width is accounted for in subaerial systems.
  Even so, variability in shear stress across a channel's width is still higher in subglacial channels, compared to subaerial ones.

  These different characteristics between glacial and subaerial channels show that records of sediment transport downstream of glaciers capture multiple subaerial processes together.
  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
  The inconsistent response of sediment transport in subglacial and subaerial systems to changing water discharge could add noise to sediment transport signals between the systems.
  In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone. 
  The differences in subaerial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.

  \section{Acknowledgments}

  I was funded by SNSF Project No. PZ00P2\_202024.
  The julia code used herein is included in the supplementary material.

\end{spacing}

\bibliographystyle{apalike}
\bibliography{Paperlib.bib}

\section{Supplement}

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

% LocalWords:  pre todo phillips CVODE decadal refreezing Mauro eq%  LocalWords:  dQs
% LocalWords:  dx ms ca DF Hm meltwater Griessee gridded ehrbar ng
% LocalWords:  koppes costa bendixen delaney herman hallet creyts ian
% LocalWords:  farinotti stott rickenman delaney2019 plagerized
% LocalWords:  fischer zekollari clarke raymond larour delaneyinprep
% LocalWords:  guillon exner wolman engelund Weisbach