### 2025-02-03 revised_v1/response_editor.tex
small changes

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}
Dear Editor,

\vspace{.3cm}

We thank you for taking on this manuscript and coordinating the reviews. We found both reviews very helpful for improving the manuscripts.

\vspace{.2cm}

Following Reviewer 1, we have modified the first objective of the manuscript to describe the conditions where covariance between sediment transport capacity and water discharge could occur.
To develop this point further, we have conducted additional experiments with a series of smoothed hydrographs to better quantify the conditions when covariance could occur.
Additionally, we have reorganized and modified the text in response to their comments.
Lastly, at several points, in the review, it was mentioned that a point was already established.
It could be a substantial oversight of the literature on our part as authors.
However, we would be grateful if the reviewer would provide the specific references in these cases to help us fully address their comments.

\vspace{.2cm}

With respect to Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in questions.
We understand their critique about the accumulation of water flow along the glacier's length.
However, we cannot find a rational that would suggest that these findings would not over shorter channels or in a more complex (i.e. spatially distributed) model.
To further highlight their concerns, we have moved the ``Experiment limitations'' discussion to its own section following the ``Results.''

\vspace{.2cm}
Again we are grateful for your assistance with the manuscript and hope that you find our responses compelling

\vspace{.5cm}

Best regards,

\vspace{.75cm}

Ian Delaney on behalf of all authors

\bibliographystyle{apalike} 
\bibliography{../PaperLib.bib}
\end{document}