### 2025-02-03 revised_v1/response_editor.tex

\documentclass[11pt]{article}
\usepackage{lscape}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\setlength{\parindent}{0pt}
\usepackage [colon]{natbib}
\usepackage {comment}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[a4paper, margin=2.40cm]{geometry}
\usepackage{doi}
\usepackage{multirow}
\usepackage[T1, OT1]{fontenc}
\DeclareTextSymbolDefault{\dh}{T1}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial
\usepackage[modulo]{lineno}
\usepackage{wrapfig}
% \linenumbers
\usepackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{amsmath}
\usepackage{float}
% \errorcontextlines=3
\usepackage[explicit]{titlesec}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[export]{adjustbox}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{bibentry}
\usepackage{tikz}
\usepackage[inline,shortlabels]{enumitem}
\usepackage[all]{nowidow}

\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\leif}[1]{{\textbf{\color{red}Leif says:} \color{red} #1} }
\newcommand{\frederic}[1]{{\textbf{\color{red}Frederic says:} \color{red} #1} }
\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\title{}
\author{}

\begin{document}
Dear Editor,

We thank you for handling this manuscript and coordinating the reviews. We found both reviews very helpful in improving the manuscript.

\vspace{.2cm}

Following Reviewer 1’s feedback, we have revised the first objective of the manuscript to better describe the conditions under which covariance between sediment transport capacity and water discharge may occur. To further develop this point, we conducted additional experiments using a series of smoothed hydrographs to better quantify the conditions that lead to covariance. Additionally, we have reorganized and revised the text in response to their comments.

At several points in the review, it was noted that a particular point had already been established. If this reflects a substantial oversight of the literature on our part, we sincerely appreciate the reviewer bringing it to our attention. However, we would be grateful if the reviewer could provide specific references in these instances to help us fully address their concerns.

\vspace{.2cm}

Regarding Reviewer 2, we believe that their specific comments about the model have been addressed by generalizing the equations in question. We acknowledge their critique concerning the accumulation of water flow along the glacier’s length. However, we do not see a rationale suggesting that these findings would not hold for shorter channels or within a more complex (i.e., spatially distributed) model. To further emphasize these concerns, we have moved the ``Experiment Limitations'' discussion to a separate section following the ``Results'' section.

\vspace{.2cm}

Once again, we sincerely appreciate your assistance with this manuscript and hope that you find our responses satisfactory.

\vspace{.5cm}

Best regards,

\vspace{.75cm}

Ian Delaney on behalf of all authors

\bibliographystyle{apalike} 
\bibliography{../PaperLib.bib}
\end{document}